\SW: II. New lens candidates from the CFHTLS discovered through citizen science

### 2015-04-19 doc/sw-cfhtls.tex

In \Fref{fig:remudist}, we show the Einstein radius distribution for the
galaxy-scale and groups-scale simulated lenses. For comparison, we give
the expected distributions (blue solid curves) for an SIS like density
profile at galaxy-scales and an NFW+Hernquist profile at groups-scales.
The theoretical curves are taken from \citet{More2012} wherein the
models are explained in detail. We note that the model we adopt at
groups-scale also includes SIS or SIE components for the group members
unlike the theoretical prediction. The theoretical curves have arbitrary
normalizations. We also show the distribution of the total magnification
for all three samples.

@@

We use the publicly available redshifts for the lens galaxy from the
\cfhtls photometric catalogs \citep{Coupon2009}. The total flux of the
lensed image or arc is measured in the $g$ band but the adopted method
is different for different samples. For the simulated sample, we
multiply the magnification of the second brightest image with the source
magnitude. For the \rf sample, the arcs are detected in the scaled
difference image of $g$ and $i$ bands from which the lensing galaxy is
subtracted \citep[for details, see]{Gavazzi2014}. Here, we use the flux
of the lensed images measured by {\sc sextractor} from the scaled
difference image, that is, $g-\alpha i$ and convert it to the $g$ band
flux using mean colors of the foreground and background population. For
the \af and the \sw sample, we integrate the flux in the image pixels
identified by \af or {\sc sextractor}.

The Einstein radius ($R_{\rm E}$) is also measured differently for
different samples. For the galaxy-scale lenses in the simulated sample,
we use the value of the input parameter of the lens model for the $R_{\rm E}$.
For groups-scale lenses, since the lens model is multi-component, we
need to determine the $R_{\rm E}$ from the image positions. We use those
pairs of lensed images that have the smallest and the largest angular
separations. The $R_{\rm E}$ here is then half of the averaged values of
these angular image separations. For the \rf sample, we use the peak
position of the lensed images measured by running {\sc sextractor} on
the scaled difference image. We calculate the image separation from the
lens center as an estimate of the $R_{\rm E}$. For the \af (SARCS)
sample, we use the same definition as above except that the peak
position is identified either by the \af or manually. For the \sw lens
sample, the same definition is used where the peak positions are
identified either with \af or {\sc sextractor}.

@@

We determine what fraction of the known sample of lenses are recovered
by \sw. Note that this sample corresponds to the \rf and \af samples
combined, as mentioned in \Sref{sec:data:kls}. In Table~\ref{tab:stats},
we show that $\sim$~25\%-30\% of the known lens candidates and
$\sim$~60\% of the known lenses are found at \StageOne and \StageTwo in
\sw. The left and the middle panels of \Fref{fig:compre} show the
fraction of detections as a function of arc magnitude and the Einstein
radius of the lens systems for the known confirmed lenses and lens
candidates. As expected, we find that systems with brighter images
and/or with larger Einstein radii are detected more often in \sw.

We find that most of the confirmed lenses and candidates that are missed
by \sw are systems with fainter arcs and smaller Einstein radii and they
come from the \rf sample. The main reason why \rf found such candidates
is because their team used lensing galaxy-subtracted images to detect
the presence of the lensed images both during the automated
object-finding phase and during the visual inspection and classification
of their candidates. This approach naturally improves the detection
efficiency at smaller Einstein radii and for fainter systems. The \sw
volunteers were not shown any galaxy-subtracted images. However, showing
galaxy subtracted images might be a better strategy to adopt for future
lens searches at galaxy-scales with \sw. In the Discussion
\Sref{sec:fn}, we further explore and discuss why some of the confirmed
lenses are missed by \sw.

@@

The distribution of image separations (i.e. twice the Einstein radius)
can be used to probe the average density profile of the lens population
\citep{Oguri2006,More2012}.  However, the lens sample found by the \af
may have incompleteness as a function of the image separation. Thus, the
lack of understanding of the selection function of the lens sample may
affect the constraints on the density profile. A blind lens search done
by visual inspection alone, for example, through \sw citizen scientists may find
lenses missed by the \af search and thereby, improve completeness.

Indeed, we have found several new lens candidates that were not known
before. In \Fref{fig:isd}, we show the image separation distribution for
the known \rf and \af sample (green), for the \sw identified (known and
new) lens sample (blue) only and combined \cfhtls sample of \rf, \af and
the new \sw lens sample (magenta). It is interesting to note that both
these samples have very similar profiles and thus, the profile of the
combined sample has not changed signficantly. This implies that previous
constraints on the image separation distribution are robust and the \af
selected sample does not suffer from significant incompleteness for
medium to large Einstein radii. This is the regime that probes density
profiles of galaxy groups to clusters.

@@

\item Compared to the sample of robotically (the \rf and \af) detected
lens candidates, the \sw sample finds lens systems with similar range in
the lens redshift and properties of the lensed images such as the total
magnitude and the Einstein radii. However, the \sw search being a purely
visual one cannot find some of the \rf-identified-lensed images with
sub-arcsecond Einstein radii because the flux of the typically faint
lensed images is obscured by the flux from bright lensing galaxies. The
\rf finds these candidates because it works with lensing
galaxy-subtracted images.

@@

SW ID & Name & RA & Dec &  $z_{\rm phot}$ & $m_i$ & $R_{\rm A}$ & G & ZooID & P & Comments  \\

@@

 SW7 & CFHTLS J220256+023432 &  330.7369 &    2.5758 &  0.0 &  0.0 &  6.8 &  2.7 & ASW0007e08 &  0.8  &  A,G/C   \\

@@

SW41 & CFHTLS J221716+015826 &  334.3189 &    1.9739 &  0.1 & 21.6 &  1.0 &  1.7 & ASW00096rm &  1.0  &  A/R,R   \\
SW48 & CFHTLS J090319-040146 &  135.8311 &   -4.0297 &  0.0 & 19.8 &  1.2 &  1.3 & ASW00007ls &  0.5  &  A,R/E   \\

@@

SW50 & CFHTLS J135724+561614 &  209.3536 &   56.2707 &  0.0 &  0.0 &  2.6 &  1.3 & ASW0006e0o &  0.9  &  D,E   \\

@@

\end{longtable}
\end{center}

>>>>

In \Fref{fig:remudist}, we show the Einstein radius ($R_{\rm E}$) distribution for the
galaxy-scale and groups-scale simulated lenses. For comparison, we give
the expected distributions (blue solid curves) for an SIS like density
profile at galaxy-scales and an NFW+Hernquist profile at groups-scales.
The theoretical curves are taken from \citet{More2012} wherein the
models are explained in detail. We note that the model we adopt at
groups-scale also includes SIS or SIE components for the group members
unlike the theoretical prediction. The theoretical curves have arbitrary
normalizations. We also show the distribution of the total magnification
for all three samples.

@@

We use the publicly available redshifts for the lens galaxy from the
\cfhtls photometric catalogs \citep{Coupon2009}. The total flux of the
lensed image or arc is measured in the $g$ band but the adopted method
is different for different samples. For the simulated sample, we
multiply the magnification of the second brightest image with the source
magnitude. For the \rf sample, the arcs are detected in the scaled
difference image of $g$ and $i$ bands from which the lensing galaxy is
subtracted \citep[for details, see][]{Gavazzi2014}. Here, we use the flux
of the lensed images measured by {\sc sextractor} from the scaled
difference image, that is, $g-\alpha i$ and convert it to the $g$ band
flux using mean colors of the foreground and background population. For
the \af and the \sw sample, we integrate the flux in the image pixels
identified by \af or {\sc sextractor}.

The Einstein radius is also measured differently for
different samples. For the galaxy-scale lenses in the simulated sample,
we use the value of the input parameter of the lens model for the $R_{\rm E}$.
For groups-scale lenses, since the lens model is multi-component, we
need to determine the $R_{\rm E}$ from the image positions. We use those
pairs of lensed images that have the smallest and the largest angular
separations. The $R_{\rm E}$ here is then half of the averaged values of
these angular image separations. For the \rf sample, we use the peak
position of the lensed images measured by running {\sc sextractor} on
the scaled difference image. We calculate the image separation from the
lens center as an estimate of the $R_{\rm E}$. For the \af (SARCS)
sample, we use the same definition as above except that the peak
position is identified either by the \af or manually. For the \sw lens
sample, the same definition is used where the peak positions are
identified either with \af or {\sc sextractor}.

@@

We determine what fraction of the known sample of lenses are recovered
by \sw. Note that this sample corresponds to the \rf and \af samples
combined, as mentioned in \Sref{sec:data:kls}. In Table~\ref{tab:stats},
we show that $\sim$~25\%-30\% of the known lens candidates and
$\sim$~60\% of the known lenses are found at \StageOne and \StageTwo in
\sw. The left and the middle panels of \Fref{fig:compre} show the
fraction of detections as a function of arc magnitude and the Einstein
radius of the lens systems for the known confirmed lenses and lens
candidates. As expected, we find that systems with brighter images
and/or with larger $R_{\rm E}$ are detected more often in \sw.

We find that most of the confirmed lenses and candidates that are missed
by \sw are systems with fainter arcs and smaller $R_{\rm E}$ and they
come from the \rf sample. The main reason why \rf found such candidates
is because their team used lensing galaxy-subtracted images to detect
the presence of the lensed images both during the automated
object-finding phase and during the visual inspection and classification
of their candidates. This approach naturally improves the detection
efficiency at smaller $R_{\rm E}$ and for fainter systems. The \sw
volunteers were not shown any galaxy-subtracted images. However, showing
galaxy subtracted images might be a better strategy to adopt for future
lens searches at galaxy-scales with \sw. In the Discussion
\Sref{sec:fn}, we further explore and discuss why some of the confirmed
lenses are missed by \sw.

@@

The distribution of image separations (i.e. 2~$R_{\rm E}$)
can be used to probe the average density profile of the lens population
\citep{Oguri2006,More2012}.  However, the lens sample found by the \af
may have incompleteness as a function of the image separation. Thus, the
lack of understanding of the selection function of the lens sample may
affect the constraints on the density profile. A blind lens search done
by visual inspection alone, for example, through \sw citizen scientists may find
lenses missed by the \af search and thereby, improve completeness.

Indeed, we have found several new lens candidates that were not known
before. In \Fref{fig:isd}, we show the image separation distribution for
the known \rf and \af sample (green), for the \sw identified (known and
new) lens sample (blue) only and combined \cfhtls sample of \rf, \af and
the new \sw lens sample (magenta). It is interesting to note that both
these samples have very similar profiles and thus, the profile of the
combined sample has not changed signficantly. This implies that previous
constraints on the image separation distribution are robust and the \af
selected sample does not suffer from significant incompleteness for
medium to large $R_{\rm E}$. This is the regime that probes density
profiles of galaxy groups to clusters.

@@

\item Compared to the sample of robotically (the \rf and \af) detected
lens candidates, the \sw sample finds lens systems with similar range in
the lens redshift and properties of the lensed images such as the total
magnitude and the Einstein radii. However, the \sw search being a purely
visual one cannot find some of the \rf-identified-lensed images with
sub-arcsecond $R_{\rm E}$ because the flux of the typically faint
lensed images is obscured by the flux from bright lensing galaxies. The
\rf finds these candidates because it works with lensing
galaxy-subtracted images.

@@

SW ID & Name & RA & Dec &  $z_{\rm phot}$ & $m_i$ & $R_{\rm E}$ & G & ZooID & P & Comments  \\

@@

@@

SW48 & CFHTLS J090319-040146 &  135.8311 &   -4.0297 &  0.0 & 19.8 &  1.2 &  1.3 & ASW00007ls &  0.5  &  A,R/E/nbr   \\ 

@@

SW50 & CFHTLS J135724+561614 &  209.3536 &   56.2707 &  0.0 &  0.0 &  2.6 &  1.3 & ASW0006e0o &  0.9  &  D,E/unlikely   \\ 

@@

 SW7 & CFHTLS J220256+023432 &  330.7369 &    2.5758 &  0.0 &  0.0 &  6.8 &  2.7 & ASW0007e08 &  0.8  &  A,G/Clus   \\ 
SW41 & CFHTLS J221716+015826 &  334.3189 &    1.9739 &  0.1 & 21.6 &  1.0 &  1.7 & ASW00096rm &  1.0  &  A/R,R/unlikely   \\ 
\end{longtable}
\end{center}
